---
title: "ECON 3003 – Homework #1 Solutions"
author: "Sarah"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    css: custom_styles.css
    includes:
      in_header: mermaid_header.html
  pdf_document:
    toc: false
    number_sections: true
    latex_engine: xelatex
header-includes:
  - \usepackage{amsmath}
  - \usepackage{booktabs}
  - \usepackage{array}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)
```

\newpage

# Question 1 [7 pts] – Extensive Form Game

**Strategic situation:** Javonie guesses which door Lycenia is behind; Malik (host) chooses a color; Javonie picks a door.

**Players:** Lycenia (L), Malik (M), Javonie (J)

**Order of play:**

1. Lycenia chooses A or B (Javonie not present)
2. Malik observes Lycenia's choice and chooses Red or Green
3. Javonie observes Malik's color (but not Lycenia's choice) and chooses A or B

**Payoffs:** (Javonie, Lycenia, Malik)

- Javonie: \$100 if correct door, \$0 if wrong
- Lycenia: \$100 if Javonie wrong, \$0 if correct
- Malik: 10 if Javonie picks A, 0 if Javonie picks B

**Extensive form (decision tree):**

<div class="mermaid">
flowchart TB
    L[Lycenia] --> A[Door A]
    L --> B[Door B]
    A --> M1[Malik]
    B --> M2[Malik]
    M1 --> R1[Red]
    M1 --> G1[Green]
    M2 --> R2[Red]
    M2 --> G2[Green]
    R1 --> J1[Javonie]
    G1 --> J2[Javonie]
    R2 --> J3[Javonie]
    G2 --> J4[Javonie]
    J1 -->|A| P1["(100,0,10)"]
    J1 -->|B| P2["(0,100,0)"]
    J2 -->|A| P3["(100,0,10)"]
    J2 -->|B| P4["(0,100,0)"]
    J3 -->|A| P5["(0,100,10)"]
    J3 -->|B| P6["(100,0,0)"]
    J4 -->|A| P7["(0,100,10)"]
    J4 -->|B| P8["(100,0,0)"]
</div>

**Text version (if diagram doesn't render):**

```
                    Lycenia
                   /      \
              Door A      Door B
                 |           |
               Malik       Malik
              /     \     /     \
           Red   Green Red   Green
             |     |     |     |
          Javonie (info sets: sees Red vs sees Green)
             |     |     |     |
          A   B   A   B  ...  (terminal payoffs below)
```

*(Information sets: Javonie's nodes after Red form one set; her nodes after Green form another. She knows the color but not Lycenia's door.)*

**Terminal payoffs (J, L, M):**

- Lycenia A, Malik Red: J picks A $\rightarrow$ (100, 0, 10); J picks B $\rightarrow$ (0, 100, 0)
- Lycenia A, Malik Green: J picks A $\rightarrow$ (100, 0, 10); J picks B $\rightarrow$ (0, 100, 0)
- Lycenia B, Malik Red: J picks A $\rightarrow$ (0, 100, 10); J picks B $\rightarrow$ (100, 0, 0)
- Lycenia B, Malik Green: J picks A $\rightarrow$ (0, 100, 10); J picks B $\rightarrow$ (100, 0, 0)

**Information sets:**

- Lycenia: one decision node (no information set)
- Malik: two nodes (knows Lycenia's choice)
- Javonie: two information sets — one when Malik chose Red, one when Malik chose Green (she does not know Lycenia's choice)

\newpage

# Question 2 [8 pts] – Normal Form from Extensive Form

*Note: The extensive-form diagram is in the PDF. The method is:*

1. Identify each player's strategies (complete plans of action for every node/information set)
2. Build the normal-form matrix with strategies as rows/columns
3. Enter payoffs for each strategy profile

**General procedure:**

- For each information set, list the actions
- A strategy for a player is a function from each of their information sets to an action
- Enumerate all such strategies and fill the payoff matrix

*(Apply this to the specific tree in the assignment.)*

\newpage

# Question 3 [6 pts total]

**Game:**

```{r q3-table, results='asis'}
q3 <- data.frame(
  Row = c("U", "D", "Payoff"),
  L = c("3, 3", "4, 1", "(P1, P2)"),
  R = c("2, 0", "8, -1", "")
)
kable(q3, col.names = c("", "L", "R"), align = c("l", "c", "c"), booktabs = TRUE)
```

## (i) Strictly and weakly dominated strategies [2]

**Player 2:**
- If P1 plays U: L gives 3, R gives 0 $\rightarrow$ L $>$ R
- If P1 plays D: L gives 1, R gives $-1$ $\rightarrow$ L $>$ R

So **R is strictly dominated by L** for Player 2.

**Player 1:**
- If P2 plays L: U gives 3, D gives 4 $\rightarrow$ D $>$ U
- If P2 plays R: U gives 2, D gives 8 $\rightarrow$ D $>$ U

So **U is strictly dominated by D** for Player 1.

**Weakly dominated:** None.

## (ii) Pure strategy Nash equilibrium [2]

After removing dominated strategies, only **(D, L)** remains.

**Check:**
- Given P2 plays L, P1's best reply is D (4 $>$ 3)
- Given P1 plays D, P2's best reply is L (1 $>$ $-1$)

**Nash equilibrium: (D, L) with payoff (4, 1).**

## (iii) Dominant strategy equilibrium? [2]

Yes. **D** is a dominant strategy for Player 1, and **L** is a dominant strategy for Player 2. So **(D, L)** is a **dominant strategy equilibrium** as well as Nash.

\newpage

# Question 4 [8 pts total]

**Game:**

```{r q4-table, results='asis'}
q4 <- data.frame(
  Row = c("U", "M", "D", "Payoff"),
  L = c("5, 9", "3, 2", "2, 8", "(P1, P2)"),
  C = c("0, 1", "0, 9", "0, 1", ""),
  R = c("4, 3", "1, 1", "8, 4", "")
)
kable(q4, col.names = c("", "L", "C", "R"), align = c("l", "c", "c", "c"), booktabs = TRUE)
```

## (i) Weakly and strictly dominated strategies [2]

**Player 2:** R is strictly dominated by L (L $\geq$ R in every row, strict when P1 plays U or D).

**Player 1:** No strategy strictly dominated in the full game. After removing R, M and D are strictly dominated by U when P2 plays L.

## (ii) Pure strategy Nash equilibria [4]

**Best responses:**
- BR$_1$(L) = U; BR$_1$(C) = U, M, D; BR$_1$(R) = D
- BR$_2$(U) = L; BR$_2$(M) = C; BR$_2$(D) = L

**Mutual best replies:**
- (U, L): BR$_1$(L) = U, BR$_2$(U) = L $\checkmark$
- (M, C): BR$_1$(C) includes M, BR$_2$(M) = C $\checkmark$

**Nash equilibria: (U, L) with payoff (5, 9) and (M, C) with payoff (0, 9).**

## (iii) Iterated dominance [2]

**Iterated elimination:**
1. R strictly dominated by L for P2 $\rightarrow$ remove R
2. M and D strictly dominated by U for P1 $\rightarrow$ remove M and D
3. P2 prefers L (9 $>$ 1)

**Iterated dominance equilibrium: (U, L) with payoff (5, 9).**

So **(U, L)** can be reached by iterated dominance; **(M, C)** cannot.

\newpage

# Question 5 [16 pts total]

**Game:**

```{r q5-table, results='asis'}
q5 <- data.frame(
  Row = c("U", "D", "Payoff"),
  L = c("3, 2", "4, 3", "(P1, P2)"),
  R = c("2, 1", "1, 4", "")
)
kable(q5, col.names = c("", "L", "R"), align = c("l", "c", "c"), booktabs = TRUE)
```

## (i) Pure strategy Nash equilibrium? [4]

**Best responses:**
- BR$_1$(L) = D; BR$_1$(R) = U
- BR$_2$(U) = L; BR$_2$(D) = R

No profile is a mutual best reply. **No Nash equilibrium in pure strategies.**

## (ii) Mixed strategy Nash equilibrium [10]

Let P1 play U with probability $p$, D with $1-p$. Let P2 play L with probability $q$, R with $1-q$.

**P2 indifference:**
$$
2p + 3(1-p) = 1p + 4(1-p) \Rightarrow -p + 3 = -3p + 4 \Rightarrow 2p = 1 \Rightarrow p^* = \frac{1}{2}
$$

**P1 indifference:**
$$
3q + 2(1-q) = 4q + 1(1-q) \Rightarrow q + 2 = 3q + 1 \Rightarrow 2q = 1 \Rightarrow q^* = \frac{1}{2}
$$

**Mixed strategy Nash equilibrium:**
- P1: $(\tfrac{1}{2} U, \tfrac{1}{2} D)$
- P2: $(\tfrac{1}{2} L, \tfrac{1}{2} R)$

**Expected payoffs:** (2.5, 2.5)

## (iii) Pareto optimality [2]

Pure outcomes: (3,2), (2,1), (4,3), (1,4). The profile (4,3) Pareto dominates (2.5, 2.5). So the mixed strategy Nash equilibrium is **not Pareto optimal**.
